---
title: "eda"
author: "Brady Lamson"
format: html
editor: source
---

# Tasks

-   [x] Get basic distribution of every regressor

-   [x] Plot each regressor against deaths

-   [x] Examine possible transformations for necessary regressors

-   [ ] Look at distributions of regressors grouped by season

```{r}
library(readxl)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(fastDummies)
library(plotly)
```

```{r}
data <- readr::read_csv(file = "../data/car_accidents/accidents.csv")
data %>% head()
```

```{r}
data %>% colnames()
```

Response variable is: - deaths

Regressors are: - injuries - marijuana - alcohol - bad weather - median income - mean commuting time

# Sanity Checks

## Season

This is by definition but all seasons should be just as frequent.

```{r}
table(data$season)
```

# Univariate Analysis

## Deaths

```{r}
data %>%
    ggplot(aes(x=deaths)) +
    geom_density() +
    ggtitle("Deaths in Car Accidents") +
    labs(x="Deaths per Capita", y="Density")
```

Appears to follow a log-normal distribution, candidate for log-transformation later.

## Injuries

```{r}
fig <- data %>%
    ggplot(aes(x=injuries)) +
    # geom_histogram(color="black") +
    geom_density() +
    ggtitle("Injuries in Car Accidents") +
    labs(
        x="Injuries per 100k Residents", y="Density", 
        subtitle = "Variable seems to follow a log normal distribution."
    )

fig
ggsave(plot=fig, filename="../presentation/presentation_images/injury_dist.png")
```

## Bad Weather

```{r}
data %>%
    ggplot(aes(x=bad_weather_accidents)) +
    # geom_histogram(color="black") +
    geom_density() +
    ggtitle("Bad Weather Accidents") +
    labs(x="Accidents", y="Density")
```

## Median Income

```{r}
data %>%
    ggplot(aes(x=median_income)) +
    geom_density() +
    ggtitle("Median Income of Colorado Counties") +
    labs(x="Income ($)", y="Density")
```

## Mean Commute Time

```{r}
data %>%
    ggplot(aes(x=mean_commuting_time)) +
    geom_density() +
    ggtitle("Mean Commute Time of Colorado Residents by County") +
    labs(x="Minutes", y="Density")
```

## Takeaways

Most of these plots tend to be unimodal with a long positive tail. This isn't too surprising. We can probably assume the average accident involves 0 deaths, 0 injuries, and pretty regular weather. So a lot of my variables involves deviations from a *typical* accident.

Median income may require some normalization later just to standardize the values and make all the numbers smaller.

Of note is that many of these variables may benefit from a log-transformation due to them following a seemingly log-normal distribution.

# Bi-Variate Analysis (Death Response)

Here we'll look into how each regressor relates to the response variable, death.

## Injuries

```{r}
data %>%
    ggplot(aes(x=deaths, y=injuries, color=season)) +
    geom_point()
```

## Bad Weather

```{r}
data %>%
    ggplot(aes(x=deaths, y=bad_weather_accidents, color=season)) +
    geom_point()
```

## Median Income

```{r}
data %>%
    ggplot(aes(x=deaths, y=median_income, color=season)) +
    geom_point()
```

## Mean Commute Time

```{r}
data %>%
    ggplot(aes(x=deaths, y=mean_commuting_time, color=season)) +
    geom_point()
```

## Season

```{r}
data %>%
    ggplot(aes(x=season, y=deaths)) +
    geom_bar(stat="identity")
```

## Takeaways

There are shockingly few relationships here to speak of. I worry none of these predictors will perform particularly well, but we'll see.

# Bi-Variate Analaysis (Injury Response)

## Bad Weather

```{r}
data %>%
    ggplot(aes(x=injuries, y=bad_weather_accidents, color=season)) +
    geom_point() +
    labs(
        y="Number of accidents in bad weather",
        x="Number of injuries per 100k residents"
    )
```

## Median Income

```{r}
data %>%
    ggplot(aes(x=injuries, y=median_income, color=season)) +
    geom_point()
```

## Mean Commute Time

```{r}
data %>%
    ggplot(aes(x=injuries, y=mean_commuting_time, color=season)) +
    geom_point()
```

## Season

```{r}
data %>%
    ggplot(aes(x=season, y=injuries)) +
    geom_bar(stat="identity")
```

# Transformations

Wary about transformations due to the complications to model coefficient interpretation, but let's at least examine a few.

Here are the variables worth consideration for transformation

-   deaths
-   injuries
-   bad weather

## Deaths

```{r}
data %>%
    ggplot(aes(x=log(deaths + 1))) +
    geom_density() +
    ggtitle("Deaths in Car Accidents") +
    labs(x="Deaths per Capita", y="Density")
```

## Injuries

```{r}
fig <- data %>%
    ggplot(aes(x=log(injuries + 1))) +
    geom_density() +
    ggtitle("Injuries in Car Accidents") +
    labs(
        x="Natural Log of Injuries per 100k Residents", 
        y="Density",
        subtitle="Log transformation of injuries results in a distribution that appears more normal. "
    )

fig
ggsave(plot=fig, filename="../presentation/presentation_images/log_injury_dist.png")
```

## Bad Weather

```{r}
data %>%
    ggplot(aes(x=log(bad_weather_accidents + 1))) +
    # geom_histogram(color="black") +
    geom_density() +
    ggtitle("Bad Weather Accidents") +
    labs(x="Accidents", y="Density")
```

## Takeaways

Worth trying to use log transformed bad weather or injuries assuming they don't complicate interpretation too much. May be worth comparing models with and without the transformed variable.

# Bi-Variate Analaysis (Log-Transformed Injury Response)

## Deaths

```{r}
fig <- data %>%
    ggplot(aes(x=log(injuries + 1), y=deaths, color=season)) +
    geom_point()+
    labs(
        y="Number of deaths per 100k residents",
        x="Natural log of injuries per 100k residents"
    ) +
    ggtitle("Deaths vs. Injuries")

fig
ggsave(plot=fig, filename="../presentation/presentation_images/deaths_vs_injuries.png")
```

## Bad Weather

```{r}
fig <- data %>%
    ggplot(aes(x=log(injuries + 1), y=bad_weather_accidents, color=season)) +
    geom_point() +
    labs(
        y="Number of accidents in bad weather",
        x="Natural log of injuries per 100k residents"
    ) +
    ggtitle("Bad Weather Accidents vs. Injuries")

fig

ggsave(plot=fig, filename="../presentation/presentation_images/weather_vs_injuries.png")
```

## Median Income

```{r}
fig <- data %>%
    ggplot(aes(x=log(injuries + 1), y=median_income, color=season)) +
    geom_point()+
    labs(
        y="Median County Household Income",
        x="Natural log of injuries per 100k residents"
    ) +
    ggtitle("Household Income vs. Injuries")

fig

ggsave(plot=fig, filename="../presentation/presentation_images/income_vs_injuries.png")
```

```{r}
data %>%
    group_by(county) %>%
    summarise(injuries = mean(injuries), median_income = mean(median_income)) %>%
    ggplot(aes(x=log(injuries), y=median_income)) +
    geom_point()
```

## Mean Commute Time

```{r}
fig <- data %>%
    ggplot(aes(x=log(injuries + 1), y=mean_commuting_time, color=season)) +
    geom_point()+
    labs(
        y="Average County Commute Time",
        x="Natural log of injuries per 100k residents"
    ) +
    ggtitle("Commute Time vs. Injuries")

fig

ggsave(plot=fig, filename="../presentation/presentation_images/commute_vs_injuries.png")
```

## Season

```{r}
data %>%
    ggplot(aes(x=season, y=log(injuries + 1))) +
    geom_bar(stat="identity")
```

## Takeaways

Log transforming injuries doesn't seem to really change the relationships of these variables in meaningful ways.
