---
title: "eda"
author: "Brady Lamson"
format: html
editor: visual
---

# Tasks

-   [x] Get basic distribution of every regressor

-   [x] Plot each regressor against deaths

-   [x] Examine possible transformations for necessary regressors

-   [ ] Look at distributions of regressors grouped by season

```{r}
library(readxl)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(fastDummies)
library(plotly)
```

```{r}
data <- readr::read_csv(file = "data/car_accidents/accidents.csv")
data %>% head()
```
```{r}
data %>% colnames()
```

Response variable is:
- deaths

Regressors are:
- injuries
- marijuana
- alcohol
- bad weather
- median income
- mean commuting time

# Sanity Checks

## Season

This is by definition but all seasons should be just as frequent. 

```{r}
table(data$season)
```


# Univariate Analysis

## Deaths

```{r}
data %>%
    ggplot(aes(x=deaths)) +
    geom_density() +
    ggtitle("Deaths in Car Accidents") +
    labs(x="Deaths per Capita", y="Density")
```

Appears to follow a log-normal distribution, candidate for log-transformation later.

## Injuries

```{r}
data %>%
    ggplot(aes(x=injuries)) +
    # geom_histogram(color="black") +
    geom_density() +
    ggtitle("Injuries in Car Accidents") +
    labs(x="Injuries per Capita", y="Density")
```

## Marijuana suspected accidents

```{r}
data %>%
    ggplot(aes(x=marijuana_suspected_accidents)) +
    # geom_histogram(color="black") +
    geom_density() +
    ggtitle("Marijuana Suspected Accidents") +
    labs(x="Accidents", y="Density")
```

## Alcohol suspected accidents

```{r}
data %>%
    ggplot(aes(x=alcohol_suspected_accidents)) +
    # geom_histogram(color="black") +
    geom_density() +
    ggtitle("Alcohol Suspected Accidents") +
    labs(x="Accidents", y="Density")
```

## Bad Weather

```{r}
data %>%
    ggplot(aes(x=bad_weather_accidents)) +
    # geom_histogram(color="black") +
    geom_density() +
    ggtitle("Bad Weather Accidents") +
    labs(x="Accidents", y="Density")
```

## Median Income

```{r}
data %>%
    ggplot(aes(x=median_income)) +
    geom_density() +
    ggtitle("Median Income of Colorado Counties") +
    labs(x="Income ($)", y="Density")
```

## Mean Commute Time

```{r}
data %>%
    ggplot(aes(x=mean_commuting_time)) +
    geom_density() +
    ggtitle("Mean Commute Time of Colorado Residents by County") +
    labs(x="Minutes", y="Density")
```

## Takeaways

Most of these plots tend to be unimodal with a long positive tail. This isn't too surprising. We can probably assume the average accident involves 0 deaths, 0 injuries, regular weather and no drugs involved. So a lot of my variables involves deviations from a *typical* accident.

Median income may require some normalization later just to standardize the values and make all the numbers smaller. 

Of note is that many of these variables may benefit from a log-transformation due to them following a seemingly log-normal distribution. 

# Bi-Variate Analysis

Here we'll look into how each regressor relates to the response variable.

## Injuries

```{r}
data %>%
    ggplot(aes(x=deaths, y=injuries, color=season)) +
    geom_point()
```

## Marijuana

```{r}
data %>%
    ggplot(aes(x=deaths, y=marijuana_suspected_accidents, color=season)) +
    geom_point()
```

## Alcohol

```{r}
data %>%
    ggplot(aes(x=deaths, y=alcohol_suspected_accidents, color=season)) +
    geom_point()
```

## Bad Weather

```{r}
data %>%
    ggplot(aes(x=deaths, y=bad_weather_accidents, color=season)) +
    geom_point()
```

## Median Income

```{r}
data %>%
    ggplot(aes(x=deaths, y=median_income, color=season)) +
    geom_point()
```

## Mean Commute Time

```{r}
data %>%
    ggplot(aes(x=deaths, y=mean_commuting_time, color=season)) +
    geom_point()
```

```{r}
data %>%
    ggplot(aes(x=injuries, y=mean_commuting_time, color=season)) +
    geom_point()
```

## Season

```{r}
data %>%
    ggplot(aes(x=season, y=deaths)) +
    geom_bar(stat="identity")
```

## Takeaways

There are shockingly few relationships here to speak of. I worry none of these predictors will perform particularly well, but we'll see.

# Transformations

Wary about transformations due to the complications to model coefficient interpretation, but let's at least examine a few.

Here are the variables worth consideration for transformation

- deaths
- injuries
- marijuana
- alcohol
- bad weather

## Deaths

```{r}
data %>%
    ggplot(aes(x=log(deaths))) +
    geom_density() +
    ggtitle("Deaths in Car Accidents") +
    labs(x="Deaths per Capita", y="Density")
```

## Injuries

```{r}
data %>%
    ggplot(aes(x=log(injuries))) +
    geom_density() +
    ggtitle("Injuries in Car Accidents") +
    labs(x="Injuries per Capita", y="Density")
```

## Marijuana

```{r}
data %>%
    ggplot(aes(x=log(marijuana_suspected_accidents))) +
    # geom_histogram(color="black") +
    geom_density() +
    ggtitle("Marijuana Suspected Accidents") +
    labs(x="Accidents", y="Density")
```

## Alcohol

```{r}
data %>%
    ggplot(aes(x=log(alcohol_suspected_accidents))) +
    # geom_histogram(color="black") +
    geom_density() +
    ggtitle("Alcohol Suspected Accidents") +
    labs(x="Accidents", y="Density")
```

## Bad Weather

```{r}
data %>%
    ggplot(aes(x=log(bad_weather_accidents))) +
    # geom_histogram(color="black") +
    geom_density() +
    ggtitle("Bad Weather Accidents") +
    labs(x="Accidents", y="Density")
```

## Takeaways

Worth trying to use log transformed bad weather or injuries assuming they don't complicate interpretation too much. May be worth comparing models with and without the transformed variable.

# Checking data Quality

Skeptical of marijuana and alcohol. I have a hunch that data is only being collected in certain counties. Let's look at the original data and check.

```{r}
file_names <- paste0("data/car_accidents/CDOTRM_CD_Crash_Listing_-_", 2021:2023, ".xlsx")
og_data <- lapply(
        file_names, 
        function(file) {readxl::read_excel(file, guess_max=200000)}
    ) %>%
    dplyr::bind_rows()%>%
    select(
        "County",
        "TU-1 NM Alcohol Suspected", "TU-2 NM Alcohol Suspected", 
        "TU-1 NM Marijuana Suspected", "TU-2 NM Marijuana Suspected",
    )%>%
    rename_all(~ str_replace_all(tolower(.), "[ -]", "_"))
```

```{r}
og_data %>%
    group_by(County) %>%
    summarise(
        marijuana_counts = sum(is.na())
    )
```

```{r}
skim(og_data)
```

Yeah we have to ditch alcohol and marijuana as variables. It's built off of too incomplete of data.