---
title: "Modeling Injuries"
author: "Brady Lamson"
format: html
editor: visual
toc: true
---

# Imports

```{r}
library(dplyr)
library(ggplot2)
library(caret)
library(leaps)
library(car)
library(effects)
library(api2lm)
```

```{r}
options(digits = 7, scipen = 2)
```

```{r}
data <- readr::read_csv(file = "../data/car_accidents/accidents.csv") %>%
    mutate(season = as.factor(season))
```

# Variable Selection

## Stepwise Selection

```{r}
lmod <- lm(injuries ~ . - county - pop_per_100k, data=data)
```

```{r}
step(lmod, direction="both")
```

### Forward selection w/ BIC

Let's look at BIC to see if we'll get a smaller model candidate.

```{r}
simple_model <- lm(injuries ~ 1, data=data)
step(simple_model, scope = formula(lmod), direction = "forward", k = log(nobs(lmod)))
```

# Checking Model Structure

## Larger Model

```{r}
lm_large <- lm(
    injuries ~ season + deaths + bad_weather_accidents + median_income + mean_commuting_time,
    data=data
)

plot(lm_large$fitted.values, lm_large$residuals)
abline(h=0)
```

```{r}
qqPlot(lm_large)
```

## Simpler Model

```{r}
lm_small <- lm(
    injuries ~ season + deaths + bad_weather_accidents + mean_commuting_time,
    data=data
)

plot(lm_small$fitted.values, lm_small$residuals)
abline(h=0)
```

```{r}
qqPlot(lm_small)
```

# Log Transforming Injuries

## Variable Selection

```{r}
simple_model_log <- lm(I(log(injuries)) ~ 1, data=data)
step(simple_model_log, scope = formula(lmod), direction = "forward", k = log(nobs(lmod)))
```

## Checking Structure

```{r}
small_model_log <- lm(I(log(injuries)) ~ season + deaths + bad_weather_accidents, data=data)
```

### Fitted vs. Residuals

```{r}
plot(small_model_log$fitted.values, small_model_log$residuals)
abline(h=0)
```

### Checking Normality

```{r}
qqPlot(small_model_log)
```

# Comparing model to complete model

We'll be using cross validation for this comparison.

```{r}
red_rmse_means <- c()
full_rmse_means <- c()
set.seed(100)
cv_fold <- trainControl(method="cv", number=100)
full_model_formula = log(injuries) ~ . - county - pop_per_100k

model_small <- train(formula(small_model_log), data=data, trControl=cv_fold, method="lm")
model_full <- train(full_model_formula, data=data, trControl=cv_fold, method="lm")
resamp <- resamples(list(model_small, model_full),
                modelNames = c("reduced", "complete"))
summary_results <- summary(resamp, metric = c("RMSE", "MAE"))
summary_results
```

Tossup whether the reduced or full model has better predictive power. I will opt for the reduced model as it has comparable predictive power while not including the added complexity.

# Effect Plots

```{r}
plot(predictorEffect("deaths", small_model_log))
```

```{r}
plot(predictorEffect("bad_weather_accidents", small_model_log))
```

```{r}
plot(predictorEffect("season", small_model_log))
```

# Model Interpretation

```{r}
model_results <- summary(small_model_log)
model_results
```

## Transform Coefficients

We want to undo our log transformation here so we can actually interpret the model correctly.

```{r}
e <- exp(1)
e^model_results$coefficients[, 1]
```


## Intercept

We expect a typical Colorado county in the fall with 0 deaths and 0 bad weather accidents to see around 125.87 injuries. 

## Season

### Spring

Given the same number of deaths and bad weather accidents, we expect a county in the spring to experience about $1-0.84\approx16\%$ less injuries per capita than a county in the fall.

### Winter

Given the same number of deaths and bad weather accidents, we expect a county in the winter to experience about $1-0.70\approx30\%$ less injuries per capita than a county in the fall.

### Summer

Given the same number of deaths and bad weather accidents, we expect a county in the summer to experience $\approx20\%$ more injuries per capita than a county in the fall.

## Deaths

Given two counties in the same season with the same number of bad weather accidents per capita but one with one additional death per captia than the other, the county with more deaths will see around $0.42\%$ more injuries per capita than the other.

## Bad Weather Accidents

Given two counties in the same season with the same number of deaths per capita but one with one additional bad weather accident per captia than the other, the county with more bad weather accidents will see around $0.28\%$ more injuries per capita than the other.

# Coefficient Confidence Intervals

```{r}
confint(small_model_log)
```

